\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

The way we design and write software to do computation and AI today is poop. How poopy you ask? Hrm\dots, let me think\dots, In my approximation, if you were to use it as a fuel source, it would be able to run all the blockchain transactions across the aggregate of current and future coins for a decade.
\par
Hrm, too much? Probably. I guess you'd expect me to use sophisticated rhetoric and cite evidence to make my points. I mean, I could write something like \textit{``The imperative programming model utilized in near all of the production software produced in the last four decades has not fundamentally changed since blah blah blah..."}. I'd certainly sound more credible perhaps. Well, though I have indeed grown accustomed to writing that way, boy has it gotten old.
\par
I'm not going to do that for this book. Let's have fun. After all, Jaseci has always been more play (and art) than work. Very ambitious play granted, but play at it's core.
\par
Everything here is based on my opinion\dots no, expert \emph{ninja} opinion, and my intuition. That suffices for me, and I hope it does for you. Though I have spent decades coding and leading teams of coders and computer scientists working on the holy grail technical challenges of our time, I won't rely on that to assert credibility. ...(o\_o)... Lets let these ideas stand or die on their own merit. Its my gut that tells me that we can do better. This book describes my attempt at better. I hope you find value in it. If you do, awesome! If you don't, awesome!
