\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}

The way we design and write software to do computation and AI today sucks. How much you ask? Hrm\dots, let me think\dots, It's a vat of boiling poop, mixed with pee, slowly swirling and bubbling toward that dehydrated semi-solid state of goop that serves to repel and repulse most normal people, only attracting the few unfortunate-fortunate folks that happen to be tantalized with \gls{scat}.
\par
Hrm, too much? Probably. I guess you'd expect me to use concrete examples and cite evidence to make my points, with me being a professor and all. I mean, I could write something like \textit{``The imperative programming model utilized in near all of the production software produced in the last four decades has not fundamentally changed since blah blah blah..."} to meet expections. I'd certainly sound more credible and perhaps super smart. I have indeed grown accustomed to writing that way and boy has it gotten old. Well, I'm not going to do that here. Let's have fun. Afterall, Jaseci has never been work for me, its play. Very ambitious play granted, but play at it's core.
\par
Everything here is based on my opinion\dots no, \emph{expert} opinion, and my intution. That suffices for me, and I hope it does for you. Even though I have spent many decades coding and leading coders working on the holy grail technical challenges of our time, I won't rely on that to assert my credibility. Let these ideas stand or die on thier own merit. Its my gut that tells me that we can do better. This book describes my attempt at better. I hope you find value in it. If you do, awesome! If you don't, awesome!
