\chapter{Rants}

\section{Libraries Suck}
\label{rant:librariessuck}
\par
Because they do.
\par
Still need more reasons?
\par
Well, if you dont already know, I'm not going to tell you.
\par
\dots
\par
Still there?
\par
\dots
\par
Fine, I'll tell you.
\begin{enumerate}
    \item They suck because they create dependancies for which you must have faith in the implementer of the library to maintain and keep bug free.
    \item They suck because there are often at least 10 options to choose from with near exact features expressings slightly different idosyncratic ways.
    \item They suck because they suck.
\end{enumerate}
Don't get me wrong, we have to use libraries. I'm not saying go reimplement the wheel 15 thousand times over. But that doesn't mean they don't suck and should be avoided if possible. The best is to know your library inside and out so the moment you hit some suckitude you can pull in the library's source code into your own codebase and \gls{pwn} it as your own.

\section{Utilizing Whitespace for Scoping is Criminal (Yea, I'm looking at you Python)}
\label{rant:whitespacesucks}
\par
This whitespace debautchery perpetrated by Python and the like is one of the most perverted abuses of ASCII code 32 I've seen in computer science. It's an assult on the freedom of coders to decide the shape and structure of the beautiful sculptures their creative minds might want to actualize in syntax. Coder's fingers have a voice! And that voice deserves to be heard! The only folks that support this oppression are those in the 1\% that get paid on a per line of code basis so they can lean on these whitespace mandates to pump up their salaries at the cost of coders everywhere.
\par
``FREE THE PEOPLE! FREE THE CODE!''
\par
``FREE THE PEOPLE! FREE THE CODE!''
\par
``FREE THE PEOPLE! FREE THE CODE!''