
Jaseci is available on Github~\cite{jaseci-github} under MIT open source license and is composed of an ecosystem of tools spanning 3 packages.
These include \textbf{Jaseci Core}, its core execution engine,  \textbf{Jaseci Serv}, its diffuse runtime cloud-scale execution engine, and \textbf{Jaseci Kit}, a collection of cutting edge AI engines provided by the Jaseci community.
In addition to these main codebases, an experimental toolkit we call \textbf{Jaseci Studio} is in development to provide visual programming and debugging tooling for developers building with Jaseci.

There are a number of notable examples of Jaseci's use in production.
These users include  four selected start-up companies that have adopted Jac and Jaseci as their development engine and have already launched their products built using Jaseci.

\indent \textbf{ myca.ai}~\cite{myca-website} - a B2C personal productivity platform that uses AI to understand personal behavior trends and help users allocate their time, prioritize their tasks and achieve personal growth goals. Using Jaseci, myca.ai’s back-end development only took 1 month and myca.ai was launched within 3 months’ development to the public. Myca.ai is one of the fast growing personal growth tool and has received positive feedback from their users.

\indent \textbf{ZeroShotBot}~\cite{zsb-website} - a B2B company that develops a cutting edge conversational AI platform using Jaseci. The product development took 2 months and was done by frontend engineers. Zeroshotbot has gained significant market traction and has been in business discussions with major logos such as Volaris, Pizzahut to provide readily deployable FAQ chatbots.



\indent \textbf{Truselph}~\cite{ts-website} - A minority founded startup. Truselph creates an avatar of the person and builds conversational intelligence that allows the general public to interact with the avatar and ask questions, while the avatar will be able to provide personalized answers with emotions and facial expressions. Truselph is in partnership with Lenovo to co-develop Truselph powered Kiosks for retail stores and is in business discussions with chains such as Sephora.


\indent \textbf{ Home Lending Pal}~\cite{hlp-website} - an AI Powered Mortgage Advisor. Home Lending Pal is a minority founded start-up that helps people, especially under-served minority population to navigate through the mortgage and home purchase process. Home Lending Pal adopted Jaseci to provide two main product features: 1 - personalized mortgage advice and 2 - Kev, an AI-powered chatbot that will answer users questions about the process and give them a plan to improve their finances.